Circular Regression Trees and Forests with an Application to Probabilistic Wind Direction Forecasting

### 2019-12-06 pkg/circtree/inst/wind_paper/draft/draft.Rnw
Add info to trainingsset for linear model
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2919 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 386973da67f622605284e048ce717cb10600c9d2

To evaluate the proposed circular regression trees and forests, a case
study on
probabilistic forecasting of wind direction for two airports in Austria is
presented. In addition to the tree-based models, alternative common
approaches
are used as a benchmark in order to assess the predictive performances of
the novel methodology of circular regression trees and forests.
%Establishing a proper regression model for circular data such as animal movement
%directions, periodic measurements of neuronal activity or wind directions,
%requires specific attention regarding its representation on the unit circle.
%Commonly applied generalized linear and additive models can provide a useful toolbox, 
%however, highly non-linear link functions can complicate interpretations while
%a non-concave log-likelihood function of the usually employed von Mises distribution 
%leads to difficulties in the incorporated optimization procedure.
%Moreover, an adequate model specification can be very challenging and non-additive 
%effects might not be detected.

%Therefore, an alternative approach to modeling circular data is proposed by
%employing the von Mises distribution within the methodology of distributional 
%trees and forests. Exploiting the tree-based separation of the data set into 
%subgroups an intercept-model is fitted within each subset such that no covariates 
%are included simplifying ....
%difficulties may arise in 
%modeling smooth effects due to highly non-linear link 
%functions and increasing irregularities when moving closer to the ``origin point'' 
%(0 which is equal to $2\pi$ in the radiant case).
%%% FIXME: clear explanation of difficulties
%Therefore, we propose a new approach by applying distributional trees and 
%forests to circular data. In particular, the parameters of the employed von Mises 
%distribution are regressed on covariates by a tree-based model yielding a constant set 
%of estimated distribution parameters for each resulting subgroup, regardless of any 
%``circular shift''/the ``circular position''.
%As circular data appears in many different areas, the development of
%appropriate statistical models which can account for periodicity has become an
%important but challenging research field. For probabilistic modeling of
%circular data typically the von Mises distribution is employed. In order to
%regress its parameters on covariates, a circular regression tree is
%proposed as an alternative to more commonly-used generalized linear and additive 
%models. 
%Moreover, the resulting circular regression trees are easy to interpret, 
%can detect non-additive effects, and select covariates and their interactions automatically. 
%By combining an ensemble of circular regression trees a more stable and regularized circular
%regression forest can be built which allows for modeling smooth effects as
%well.

%To evaluate the proposed circular regression trees and forests, a case study on
%probabilistic forecasting of wind direction for two airports in Austria is
%presented. In addition to the tree-based models, alternative common approaches
%are used as a benchmark in order to assess the predictive performances of
%circular regression trees and forests. 
}
\Keywords{Distributional Trees; Random Forests; Circular Data; von Mises Distribution}

@@

    learning sample.

>>>>

To evaluate the proposed circular regression trees and forests, a case study on
probabilistic forecasting of wind direction for two airports in Austria is
presented. In addition to the tree-based models, alternative common approaches
are used as a benchmark in order to assess the predictive performances of the
novel methodology of circular regression trees and forests.
}

\Keywords{Distributional Trees; Random Forests; Circular Data; von Mises Distribution}

@@

    learning sample. To allow for seasonally varying error characteristics in both bias and slope
    coefficients, and to allow for seasonal heteroscedasticity captured by the concentration
    parameter, we use the same time-adaptive trainings approach as for the
    climatological model; hence, separate models are estimated over all observation time points,
    using the same daytime of $31$~days centered around the day of interest over
    all available years in the training data \citep{Lang+Lerch+Mayr:2019}.
